Improving migration resolution by approximating the least-squares Hessian using non-stationary amplitude and frequency matching

### 2018-05-04 chapter-conclusion/paper.tex
thesis update

\subsection{Summary}

In this thesis, I discussed several methods and applications of data matching in seismic data analysis.
Chapter 2 focuses on introducing the three data matching operators that are used in this thesis---shifting, scaling, and filtering.
Chapter 3 introduces different methods of frequency balancing using non-stationary smoothing. 
The first method to find the non-stationary smoothing {\em radius}, or number of samples each data point is averaged over with a triangle weight, took a theoretical approach based off of the assumption that the data we observe can be modeled by a summation of Ricker wavelets.
This method worked well in certain situations, but was not robust enough to work for any given data set.
In the second method, I introduced an iterative algorithm to find the smoothing radius, and this method converges quickly and works well in several presented situations. 
Finally, a modification to this algorithm was shown and allows smoothing for more complex data sets.

@@

Chapter 4 goes into more detail of the application of matching and merging high-resolution and legacy seismic images.
This example takes two seismic volumes, acquired over the same area but using different technologies, and first matches them before merging them together to produce an optimized third image.
First, the method is demonstrated on a 2D line from the Gulf of Mexico.
Then, the method is applied to a 3D seismic volume from a different part of the Gulf of Mexico.

Chapter 5 discusses another application of improving migration resolution by approximating the least-squares Hessian using non-stationary data matching operations.
An approximation to the least-squares Hessian can be calculated by solving a data matching problem between two conventionally migrated images, and the Hessian can be represented by the combination of amplitude and frequency balancing operations.
An example is applied to a 2D synthetic Sigsbee data set.

\subsection{Future work}
In the future, the work presented in chapter 5 should be extended to involve real data and 3D examples. 
It also could benefit by comparing the results of the proposed approach taken in the chapter to other previous approaches presented to approximate the least-squares Hessian \cite[]{migdec,lsamp,siklsm,pwlsrtm,diagamp,debfilt,prestack,poststack}, to see how it compares in different situations.

Another extension of this data matching procedure may be to incorporate
%Another component to data matching may incorporate 
the phase of the signal to be matched. 
Negligible improvements were made when trying to incorporate phase corrections into the high-resolution and legacy data matching problem of chapter 4, but other data matching problems could benefit from these corrections.

@@

For brevity in this thesis, code is only included for one example of the main frequency balancing algorithm presented in chapter 3.
The code for the rest of the examples in this thesis are available online at the URL above.
%The code for the rest of the examples in this thesis is available online at the URL above.

@@

\begin{table}[h]
\centering
\caption{Figures and the scripts to generate them}
\label{my-label}
\begin{tabular}{l|l|l}
\hline
        \textbf{Figures} & \textbf{Directory} & \textbf{Listings} \\ \hline 
        \ref{fig:legacy,low-freq} & chapter-locfreq/merge/ & \ref{lst:lms}, \ref{lst:lmrp}, \ref{lst:lmrc}\\ 
        \ref{fig:one0,one1,one2,one3}, \ref{fig:bef,aft} & chapter-background/dmExample/ & ---\\%\ref{lst:dm} \\ 
        \ref{fig:legacy,hires}, \ref{fig:nspectra,hires-smooth-spec}, \ref{fig:rect}, \ref{fig:freqdif,freqdif-filt}           & chapter-merge/apache/ & ---\\%\ref{lst:mas} \\ 
        \ref{fig:legacy,hires-agc}, \ref{fig:rect5}, \ref{fig:nspectra-orig,high-smooth-spec5}, \ref{fig:freqdif,freqdif-filt5}          & chapter-locfreq/merge/ &   \ref{lst:lms}, \ref{lst:lmrp}, \ref{lst:lmrc}\\ 
        \ref{fig:pp,ss}, \ref{fig:before,after}         & chapter-locfreq/vecta/&  --- \\%\ref{lst:lvs}, \ref{lst:lvn}, \ref{lst:lvrp}, \ref{lst:lvrc}\\ 
        \ref{fig:all,scalar} & chapter-locfreq/convergence/&  --- \\%\ref{lst:conv}\\ 
        \ref{fig:legacy,hires,merge2-reverse}, \ref{fig:diff0,diff1}, \ref{fig:hweight,lweight-reverse}, \ref{fig:nspectra22-reverse}           & chapter-merge/apache/ & --- \\% \ref{lst:mas} \\ 
        \ref{fig:window1,window2}, \ref{fig:nspectra2} & chapter-merge/pcable/ & --- \\%\ref{lst:mps} \\ 
        \ref{fig:legacy4,hires4,merge3}  & chapter-merge/pcable2/ &  --- \\%\ref{lst:mp2s}, \ref{lst:mp2rp}, \ref{lst:mp2rc}\\ 
        \ref{fig:mod,vel-migration}, \ref{fig:image0,image1}, \ref{fig:a0,rect10b}, \ref{fig:migdec-shap}, \ref{fig:image0-w3,migdec-w3,mod-w3} & chapter-merge/mighes/ & --- \\%\ref{lst:hss}, \ref{lst:hsrp}, \ref{lst:hsrc}\\ 
        \ref{fig:tf}, & chapter-merge/triop/ & --- \\%\ref{lst:triop}, \ref{lst:trfu} \ref{lst:solve}\\ 
\end{tabular}
\end{table}

>>>>

%\subsection{Summary}

In this thesis, I discussed several methods and applications of data matching in seismic data analysis.
Chapter \ref{ch:chapter-background} focuses on introducing the three data matching operators that are used in this thesis---shifting, scaling, and filtering.
Chapter \ref{ch:chapter-locfreq} introduces different methods of frequency balancing using non-stationary smoothing. 
The first method to find the non-stationary smoothing {\em radius}, or number of samples each data point is averaged over with a triangle weight, took a theoretical approach based off of the assumption that the data we observe can be modeled by a summation of Ricker wavelets.
This method worked well in certain situations, but was not robust enough to work for any given data set.
In the second method, I introduced an iterative algorithm to find the smoothing radius, and this method converges quickly and works well in several presented situations. 
Finally, a modification to this algorithm was shown and allows smoothing for more complex data sets.

@@

Chapter \ref{ch:chapter-merge} goes into more detail of the application of matching and merging high-resolution and legacy seismic images.
This example takes two seismic volumes, acquired over the same area but using different technologies, and first matches them before merging them together to produce an optimized third image.
First, the method is demonstrated on a 2D line from the Gulf of Mexico.
Then, the method is applied to a 3D seismic volume from a different part of the Gulf of Mexico.

Chapter \ref{ch:chapter-mighes} discusses another application of improving migration resolution by approximating the least-squares Hessian using non-stationary data matching operations.
An approximation to the least-squares Hessian can be calculated by solving a data matching problem between two conventionally migrated images, and the Hessian can be represented by the combination of amplitude and frequency balancing operations.
An example is applied to a 2D synthetic Sigsbee data set.

\subsection{Future work}
In the future, the work presented in Chapter \ref{ch:chapter-mighes} should be extended to involve real data and 3D examples. 
It also could benefit by comparing the results of the proposed approach taken in the chapter to other previous approaches presented to approximate the least-squares Hessian \cite[]{migdec,lsamp,siklsm,pwlsrtm,diagamp,debfilt,prestack,poststack}, to see how it compares in different situations.

Another extension of this data matching procedure may be to incorporate
%Another component to data matching may incorporate 
the phase of the signal to be matched. 
Negligible improvements were made when trying to incorporate phase corrections into the high-resolution and legacy data matching problem of Chapter \ref{ch:chapter-merge}, but other data matching problems could benefit from these corrections.

@@

For brevity in this thesis, code is only included for one example of the main frequency balancing algorithm presented in Chapter \ref{ch:chapter-locfreq}.
The code for the rest of the examples in this thesis are available online at the URL above.
%The code for the rest of the examples in this thesis is available online at the URL above.

@@

\begin{table}[h]
\centering
\caption{List of figures in this thesis and the locations of scripts and programs to generate them}
\label{my-label}
\begin{tabular}{l|l|l}
\hline
        \textbf{Figures} & \textbf{Directory} & \textbf{Listings} \\ \hline 
        \ref{fig:legacy,low-freq} & chapter-locfreq/merge/ & \ref{lst:lms}, \ref{lst:lmrp}, \ref{lst:lmrc}\\ 
        \ref{fig:one0,one1,one2,one3}, \ref{fig:bef,aft} & chapter-background/dmExample/ & ---\\%\ref{lst:dm} \\ 
        \ref{fig:legacy,hires}, \ref{fig:nspectra,hires-smooth-spec}, \ref{fig:rect}, \ref{fig:freqdif,freqdif-filt}           & chapter-merge/apache/ & ---\\%\ref{lst:mas} \\ 
        \ref{fig:legacy,hires-agc}, \ref{fig:rect5}, \ref{fig:nspectra-orig,high-smooth-spec5}, \ref{fig:freqdif,freqdif-filt5}          & chapter-locfreq/merge/ &   \ref{lst:lms}, \ref{lst:lmrp}, \ref{lst:lmrc}\\ 
        \ref{fig:pp,ss}, \ref{fig:before,after}         & chapter-locfreq/vecta/&  --- \\%\ref{lst:lvs}, \ref{lst:lvn}, \ref{lst:lvrp}, \ref{lst:lvrc}\\ 
        \ref{fig:all,scalar} & chapter-locfreq/convergence/&  --- \\%\ref{lst:conv}\\ 
        \ref{fig:legacy,hires,merge2-reverse}, \ref{fig:diff0,diff1}, \ref{fig:hweight,lweight-reverse}, \ref{fig:nspectra22-reverse}           & chapter-merge/apache/ & --- \\% \ref{lst:mas} \\ 
        \ref{fig:window1,window2}, \ref{fig:nspectra2} & chapter-merge/pcable/ & --- \\%\ref{lst:mps} \\ 
        \ref{fig:legacy4,hires4,merge3}  & chapter-merge/pcable2/ &  --- \\%\ref{lst:mp2s}, \ref{lst:mp2rp}, \ref{lst:mp2rc}\\ 
        \ref{fig:mod,vel-migration}, \ref{fig:image0,image1}, \ref{fig:a0,rect10b}, \ref{fig:migdec-shap}, \ref{fig:image0-w3,migdec-w3,mod-w3} & chapter-merge/mighes/ & --- \\%\ref{lst:hss}, \ref{lst:hsrp}, \ref{lst:hsrc}\\ 
        \ref{fig:tf}, & chapter-merge/triop/ & --- \\%\ref{lst:triop}, \ref{lst:trfu} \ref{lst:solve}\\ 
\end{tabular}
\end{table}

### 2018-05-04 chapter-locfreq/paper.tex
thesis almost done
thesis update

%\title{Theory}
\renewcommand{\thefootnote}{\fnsymbol{footnote}}
\title{Balancing local frequency content in seismic data using non-stationary smoothing}
\relax\footnotetext{Parts of this chapter were first published in \cite{locfreq}, \cite{merge2}, and \cite{merge}. This work was done under the supervision of Dr. Sergey Fomel.}
\author{Sarah Yvonne Greer}
\maketitle

@@

        \subsection{Iterative method for calculating the smoothing radius}
        \inputdir{merge}
        The theoretical smoothing radius works in some situations, like in the first example of the next chapter, but it does not work in situations where litte overlap is present in frequency content between data sets, like in the second example of the next chapter.
        In this case, we use an iterative method to find what the smoothing radius should be that is based on the physical understanding of what smoothing is doing.

@@

        An application of aligning the high-resolution and legacy images is discussed in the next chapter.

@@

        So far, this algorithm and its modification have been discussed when the smoothing radius is only calculated along one dimension. 
        However, Chapter 5 provides an application where the smoothing radius is calculated in multiple directions.

@@

\section{Acknowledgements}
        We thank the sponsors of the Texas Consortium for Computational Seismology (TCCS) for their financial support.

>>>>

%\title{Theory}
\renewcommand{\thefootnote}{\fnsymbol{footnote}}
\title{Balancing local frequency content in seismic data using non-stationary smoothing}
\relax\footnotetext{Parts of this chapter were first published in \cite{locfreq}, \cite{merge2}, and \cite{merge}. This work was done under the supervision of Dr. Sergey Fomel.}
\author{Sarah Yvonne Greer}
\label{ch:chapter-locfreq}

\maketitle

@@

        \subsection{Iterative method for calculating the smoothing radius}
        \inputdir{merge}
        The theoretical smoothing radius works in some situations, like in the first example of Chapter \ref{ch:chapter-merge}, but it does not work in situations where litte overlap is present in frequency content between data sets, like in the second example of Chapter \ref{ch:chapter-merge}.
        In this case, we use an iterative method to find what the smoothing radius should be that is based on the physical understanding of what smoothing is doing.

@@

        An application of aligning the high-resolution and legacy images is discussed in Chapter \ref{ch:chapter-merge}.

@@

        So far, this algorithm and its modification have been discussed when the smoothing radius is only calculated along one dimension. 
        However, Chapter \ref{ch:chapter-mighes} provides an application where the smoothing radius is calculated in multiple directions.

@@

\section{Acknowledgments}
        We thank the sponsors of the Texas Consortium for Computational Seismology (TCCS) for their financial support.

### 2018-05-04 chapter-mighes/paper.tex
thesis almost done
thesis update

\usetikzlibrary{arrows,automata}
\author{Sarah Greer*, Zhiguang Xue, and Sergey Fomel, The University of Texas at Austin}
\relax\footnotetext{Parts of this chapter were first published in \cite{mighess}. This work was done under the supervision of Dr. Sergey Fomel, and Dr. Zhiguang Xue assisted in the migration of the Sigsbee synthetic data set.}

@@

    In this paper, we take a different approach. 
    We note that the two primary differences between least squares migrated images and conventional migrated images are amplitude and frequency variations \cite[]{Hou2015, Hou2016}. 
    Here, we treat this as a data matching problem between two conventionally migrated images, and find separate operations to account for both amplitude and frequency variations.

@@

    We find the smoothing radius iteratively using the method of \cite{locfreq} from chapter 3, with a modification that allows the smoothing radius to be calculated in both spatial directions. 
Essentially, this is found by choosing an initial guess of a smoothing radius, $\mathbf{R}^{(0)}$, and updating it iteratively such that 
    \begin{equation}
        \label{eq:it}
        \mathbf{R}^{(i+1)} = \mathbf{R}^{(i)}+ \alpha \left [ \mathbf{F}[\mathbf{S}_{\mathbf{R}^{(i)}} \mathbf{m}_0] - \mathbf{F}[\mathbf{m}_1] \right ]\;,
    \end{equation}
    where $\mathbf{F}$ is the local frequency operator, $\mathbf{S}_{\mathbf{R}^{(i)}}$ is the smoothing operator of radius $\mathbf{R}$ at the $i$th iteration, and $\alpha$ is a scalar constant that represents the step length.
    After a small number of iterations, the smoothing operator is found that, once applied to $\mathbf{m}_0$, balances local frequency content with $\mathbf{m}_1$.

    For this particular application using depth migration, this operator should technically be specified to balance {\em wavenumber} instead of {\em frequency}.
    However, it is kept as frequency to keep consistent terminology with the algorithm developed in chapter 3.

@@

\section{Conclusions}
    Least-squares migration can produce an accurate migrated image, but it is more computationally expensive than conventional migration.
    In this paper, we apply an approximate inverse Hessian operator to a conventional migrated image to approximate the least-squares migrated image.
    Because the primary differences between least-squares migration and conventional migration amount to amplitude and frequency variations, we approximate the forward Hessian by calculating frequency and amplitude matching operators.
    The amplitude matching operator is found by calculating the amplitude envelopes of migrated and remigrated images and smoothly dividing them, and the frequency matching operator is found using an iterative algorithm and non-stationary smoothing.
    The Hessian is approximated by a combination of these two operators to ensure symmetry.
    This method involves a windowless approach and is cheap to calculate and apply.
    Additionally, by defining the Hessian with two separate operators, we can examine, and control, the ``ingredients'' of the Hessian operator, and see how changing them impacts the final image.

@@

\section{Acknowledgements}
    We thank the sponsors of the Texas Consortium for Computational Seismology (TCCS) for their financial support.
    The examples in this paper can be reproduced using the Madagascar open-source software package \cite[]{madagascar}.

>>>>

\usetikzlibrary{arrows,automata}
\author{Sarah Greer*, Zhiguang Xue, and Sergey Fomel, The University of Texas at Austin}
\label{ch:chapter-mighes}

\relax\footnotetext{Parts of this chapter were first published in \cite{mighess}. This work was done under the supervision of Dr. Sergey Fomel, and Dr. Zhiguang Xue assisted in the migration of the Sigsbee synthetic data set.}

@@

    In this thesis, we take a different approach. 
    We note that the two primary differences between least squares migrated images and conventional migrated images are amplitude and frequency variations \cite[]{Hou2015, Hou2016}. 
    Here, we treat this as a data matching problem between two conventionally migrated images, and find separate operations to account for both amplitude and frequency variations.

@@

    We find the smoothing radius iteratively using the method of \cite{locfreq} from Chapter \ref{ch:chapter-locfreq}, with a modification that allows the smoothing radius to be calculated in both spatial directions. 
Essentially, this is found by choosing an initial guess of a smoothing radius, $\mathbf{R}^{(0)}$, and updating it iteratively such that 
    \begin{equation}
        \label{eq:it}
        \mathbf{R}^{(i+1)} = \mathbf{R}^{(i)}+ \alpha \left [ \mathbf{F}[\mathbf{S}_{\mathbf{R}^{(i)}} \mathbf{m}_0] - \mathbf{F}[\mathbf{m}_1] \right ]\;,
    \end{equation}
    where $\mathbf{F}$ is the local frequency operator, $\mathbf{S}_{\mathbf{R}^{(i)}}$ is the smoothing operator of radius $\mathbf{R}$ at the $i$th iteration, and $\alpha$ is a scalar constant that represents the step length.
    After a small number of iterations, the smoothing operator is found that, once applied to $\mathbf{m}_0$, balances local frequency content with $\mathbf{m}_1$.

    For this particular application using depth migration, this operator should technically be specified to balance {\em wavenumber} instead of {\em frequency}.
    However, it is kept as frequency to keep consistent terminology with the algorithm developed in Chapter \ref{ch:chapter-locfreq}.

@@

\section{Conclusions}
    Least-squares migration can produce an accurate migrated image, but it is more computationally expensive than conventional migration.
    In this chapter, we apply an approximate inverse Hessian operator to a conventional migrated image to approximate the least-squares migrated image.
    Because the primary differences between least-squares migration and conventional migration amount to amplitude and frequency variations, we approximate the forward Hessian by calculating frequency and amplitude matching operators.
    The amplitude matching operator is found by calculating the amplitude envelopes of migrated and remigrated images and smoothly dividing them, and the frequency matching operator is found using an iterative algorithm and non-stationary smoothing.
    The Hessian is approximated by a combination of these two operators to ensure symmetry.
    This method involves a windowless approach and is cheap to calculate and apply.
    Additionally, by defining the Hessian with two separate operators, we can examine, and control, the ``ingredients'' of the Hessian operator, and see how changing them impacts the final image.

@@

\section{Acknowledgments}
    We thank the sponsors of the Texas Consortium for Computational Seismology (TCCS) for their financial support.
    The examples in this chapter can be reproduced using the Madagascar open-source software package \cite[]{madagascar}.
